### 2020-02-16 figures/cactus_cadical.tex
fb notes for changes

\documentclass{standalone}
\title{}
\date{}
\author{}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{tikz,pgfplotstable,pgfplots}
\usetikzlibrary{arrows}
\begin{document}

\begin{tikzpicture}
\begin{axis}[
	xlabel={Instances Solved},
	ylabel={Run Time (sec)},width=18cm,height=10cm,legend pos=north west,xmin=0,ymin=0,xmax=600,ymax=5500
]

\addplot table [x=index, y=original_time, col sep=comma] {results_cadical-master_data.csv};
\addplot table [x=index, y=original_time, col sep=comma] {results_cadical-master-chrono_data.csv};
\addplot table [x=index, y=original_time, col sep=comma] {results_cadical-master-trail-working-version_data.csv};
\addplot table [x=index, y=original_time, col sep=comma] {results_cadical-master-trail-1lvl-proper_data.csv};

\legend{cadical,cadical-chrono,cadical-trail,cadical-trail-enhanced}
\end{axis}
\end{tikzpicture}

\end{document}