### 2020-02-16 figures/test.tex

\documentclass{standalone}
\title{}
\date{}
\author{}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{tikz,pgfplotstable,pgfplots}
\usetikzlibrary{arrows}
\begin{document}

\begin{tikzpicture}
\begin{axis}[xlabel=Speedup or Slowdown Factor,ylabel=Frequency,width=12cm,height=12cm,
ybar=0pt,title=Cadical Trail Saving Run Time Comparison,
xticklabel style={/pgf/number format/fixed,/pgf/number format/precision=2},
yticklabel style={/pgf/number format/fixed,/pgf/number format/precision=5},]
\addplot table [x, y, col sep=tab] {2testdata};
\addplot table [x, y, col sep=tab] {testdata};
\legend{speedup, slowdown}
\draw [<-] (axis cs:0.1,700)-- +(10pt,10pt) node[right] {5104 instances had speedup or slowdown factor $< 1.07$};
\draw (axis cs:2.1,250) node[left] {34 instances had speedup factor $> 4$};
\draw (axis cs:2.1,200) node[left] {32 instances had slowdown factor $> 4$};
\iffalse
\draw [blue] (rel axis cs:0.54,0) -- (rel axis cs:0.54,1);
$\lvert log_2 \frac{\text{enqueue time}}{\text{trail saving time}} \rvert$
\fi
\end{axis}
\end{tikzpicture}

\end{document}