### 2020-02-16 figures/test2.tex

\documentclass{standalone}
\usepackage{tikz,pgfplotstable,pgfplots}
\pgfplotsset{compat=1.13}
\begin{document}
\begin{tikzpicture}
\begin{axis}[title=Trail Saving Running Time Per Instance,axis equal,ymin=0,xmin=0,
width=11cm,height=11cm,xlabel=original time (s),ylabel=trail saving time (s),axis x line=bottom,axis y line=left,xlabel near ticks,ylabel near ticks]
\addplot[only marks,mark size=1pt,mark=*,mark options={fill=black}]
	table[x=original,y=trail]
{test2.txt};
\addplot [red,samples at={0,1}] {x};
\draw [blue,dashed] (rel axis cs:0,0) -- (rel axis cs:1,1);
\end{axis}
\end{tikzpicture}
\end{document}