### 2020-02-28 figures/cactus_cadical_new.tex
plot

\documentclass{standalone}
\title{}
\date{}
\author{}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{tikz,pgfplotstable,pgfplots}
\usetikzlibrary{arrows}
\begin{document}

\begin{tikzpicture}
\begin{axis}[
	xlabel={Instances Solved},
	ylabel={Run Time (sec)},width=18cm,height=12cm,legend pos=north west,xmin=400,ymin=1000,xmax=540,ymax=5100
]

\addplot[red,mark=x] table [x=index, y=original_time, col sep=comma] {results_cadical_nochrono_data.csv};
\addplot[green,mark=o] table [x=index, y=original_time, col sep=comma] {results_cadical_chrono_data.csv};
\addplot[blue,mark=+] table [x=index, y=original_time, col sep=comma] {results_cadical_nochrono_trail_data.csv};
\addplot[cyan,mark=square] table [x=index, y=original_time, col sep=comma] {results_cadical_nochrono_trail_1lvl_data.csv};
\addplot[gray,mark=triangle] table [x=index, y=original_time, col sep=comma] {results_cadical_nochrono_trail_1lvl_10reason_data.csv};

\legend{cadical,cadical-chrono,cadical-trail,cadical-trail-multipleBT-lookahead, cadical-trail-multipleBT-lookahead-reason}
\end{axis}
\end{tikzpicture}

\end{document}